\section{Limitations}
\label{sec:limitations}

Our eight models share the 2024--2025 instruction-tuning paradigm; temporal correlation may limit exchangeability, and results do not generalize to API-based or base models. Preliminary scale validation (\S\ref{sec:scale_validation}) with Qwen2.5 (3B/7B/72B) and cross-family Llama-3.1-70B confirms consistent variance structure across two architecture families, but validation beyond these two families is needed before extrapolating budget recommendations to arbitrary architectures.

The benchmark sample ($n = 5$) limits statistical power for format-associated observations ($p = 0.10$); other formats (code generation, multi-turn dialogue) may differ. All D-study budgets reported in this paper are strictly valid only under binary correctness scoring. Our partial-credit analysis (GSM8K, 8 models) shows $G$ drops from 0.86 to 0.77 (crossed design) with all models showing $G_\text{partial} < G_\text{binary}$, suggesting limited transferability to finer-grained scoring rubrics; the analysis is limited to a single benchmark, and its generalizability to other tasks is unknown. Our temperature range (0.0--0.7) does not cover high-temperature regimes ($\tau > 1.0$) sometimes used in generation tasks. Our design varies few-shot example ordering but not example identity, a known variance source. Results depend on vLLM~0.9.2; different inference engines or vLLM versions may alter CUDA-level nondeterminism patterns.

All eight models may have encountered benchmark items during pre-training; differential contamination across architectures could inflate model$\times$item interaction \citep{golchin2023contamination, oren2024proving}, though consistent variance structures across three independently constructed MC benchmarks provide indirect evidence against contamination as the sole driver. Contamination affects absolute budget estimates but has limited impact on the multi-facet vs.\ single-condition budget \emph{ratio}, since both designs are equally exposed to contaminated items.