\section{Related Work}
\label{sec:related}

\paragraph{Variance Decomposition in ML/NLP Evaluation.}
Factorial decomposition of evaluation variance has roots in the ML reproducibility literature. \citet{bouthillier2021accounting} introduced an ANOVA-based template that decomposes benchmark variance across implementation details, hyperparameters, and data splits, establishing that design choices account for a substantial share of reproducibility failures. Subsequent studies applied similar decomposition to specific evaluation facets \citep{dodge2020finetuning, henderson2018deep, magnusson2023reproducibility}. \citet{messing2026hidden} brought Generalizability theory into LLM evaluation, decomposing variance across five methodology facets on a single benchmark (judge identity: 43.8\%; see Positioning below for detailed comparison). \citet{madaan2024quantifying} proposed per-factor variance metrics for LLM evaluation benchmarks but did not cross factors simultaneously. Holistic evaluation frameworks \citep{liang2023helm, kiela2021dynabench} standardize multi-scenario comparison but do not decompose variance sources. \citet{blackwell2024towards} quantified uncertainty in benchmark scores, and \citet{qian2026benchmark} conducted systematic meta-evaluation of LLM benchmarks; \citet{burnell2023rethink} called for rethinking evaluation reporting standards; \citet{saxon2024metrology} advocated model metrology as a framework for rigorous benchmark design. Recently, G-theory has been applied to LLM-based writing assessment \citep{song2025gtheory} and annotation reliability \citep{varannot2026gtheory}, but these studies address scoring consistency rather than benchmark evaluation variance. \citet{kunievsky2025intent} proposed an alternative decomposition that separates intent uncertainty, articulation variance, and model uncertainty, but this framework does not cross infrastructure and methodology facets in a factorial design. Our work extends these approaches by crossing seven facets across five benchmarks, enabling direct comparison of variance structures across evaluation formats.

\paragraph{Infrastructure and Prompt Variance.}
A parallel line of work examines individual variance sources in isolation. \citet{yuan2025layercast} documented up to 9\% accuracy variation from numerical precision; \citet{messina2026background} and \citet{thinkingmachines2025batch} formalized and mitigated CUDA nondeterminism from floating-point non-associativity. On the methodology side, \citet{mizrahi2024state} advocated multi-prompt evaluation based on large-scale sensitivity analysis; \citet{sclar2024formatting} and BrittleBench \citep{brittlebench2026} found 50--76\% accuracy swings from prompt formatting; \citet{hua2025flaw} showed that much of this sensitivity is an artifact of log-likelihood scoring (our generation-based protocol controls for this). \citet{pezeshkpour2024ordering} and \citet{alzahrani2024benchmarks} documented ordering and benchmark sensitivity; \citet{bui2025seeds} and \citet{biderman2024lessons} documented seed and training sensitivity respectively. Each isolates one factor, precluding interaction estimation. Standardized evaluation frameworks such as lm-evaluation-harness \citep{gao2024lmeval} reduce implementation variance but do not control prompt, seed, or ordering facets.

\paragraph{IRT and Efficient Benchmarking.}
Item Response Theory (IRT) offers a complementary perspective on evaluation design by modeling item difficulty and discrimination parameters. \citet{lalor2019irt} introduced IRT-based item analysis for NLP benchmarks, and \citet{rodriguez2021leaderboards} showed that IRT reveals hidden properties of evaluation items that raw accuracy scores obscure. Building on this foundation, \citet{polo2024tinybenchmarks} developed tinyBenchmarks, using IRT-based subsampling to reduce evaluation cost while preserving ranking accuracy, and \citet{zhou2026irt} recommended a 116-item subset via PSN-IRT for efficient MMLU evaluation. \citet{perlitz2023efficient} and \citet{kipnis2025metabench} explored related strategies for benchmark compression; \citet{li2025atlas} extended IRT to adaptive testing. Complementary benchmark revision efforts address item quality directly: MMLU-Redux \citep{gema2025mmlu} identified annotation errors in MMLU, while MMLU-Pro \citep{wang2024mmlupro} extended it with harder items and expanded option sets. These approaches optimize item selection or quality but do not model variance from prompt, seed, or ordering facets. IRT-selected items evaluated under a single methodological condition can therefore systematically underestimate the number of items required for reliable measurement across diverse evaluation settings.

\paragraph{Psychometrics Foundations.}
G-theory \cite{shavelson1991generalizability, brennan2001generalizability, cardinet2010applying} provides the canonical framework for multi-facet reliability analysis; \citet{webb2006reliability} detail the random versus fixed facet distinctions central to our model specification. Extensions to binary data include \citet{mushquash2006gtheory} and logistic-link generalized G-theory \cite{jiang2018bayesian, tuerlinckx2006glmm}. We operationalize differential item functioning---established as a formal framework by \citet{holland1993dif} and operationalized via statistical detection procedures by \citet{clauser1998dif}---as model$\times$item interaction. Recent psychometric perspectives on LLM evaluation address construct validity \cite{kearns2026construct, bean2025measuring, freiesleben2026construct} and measurement properties \cite{ye2025psychometrics}; \citet{zhu2026benchmark} propose a benchmark health index for systematic meta-evaluation.

\paragraph{Positioning.}
Closest to our work, \citet{messing2026hidden} applied G-theory to LLM evaluation through their TEE framework, decomposing five methodology facets on a single benchmark (judge identity: 43.8\%). We extend this in three directions: (i) we fix the judge to the model itself, replacing judge identity with infrastructure facets and additional methodology facets (seed, ordering); (ii) we cross facets across five benchmarks, enabling cross-format comparison; and (iii) we translate variance components into reproducibility budgets via D-study projections. The two studies are complementary: TEE characterizes judge selection, while we characterize item-and-protocol selection. Relative to IRT approaches \citep{polo2024tinybenchmarks, zhou2026irt}, G-theory optimizes an orthogonal dimension: conditions needed for target reliability rather than item discrimination.